Markowitz Portfolio Optimization with Inflation

### 2023-12-18 paper/paper.tex
[LATEX] Fix compilation errors in paper and presentation

\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage[left=3cm, right=3cm, bottom=3cm, top=3cm]{geometry}	
\usepackage{setspace}
\onehalfspacing
\usepackage{float}
\usepackage{amsmath}

\title{Markowitz Portfolio Optimization with Inflation}
\author{Pablo Duce Cabeza, Djordje Nikolic, Sergey Mirzoev, Marc Tschudi}
\date{December 2023}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{figure/uzh_logo_e_pos.eps}
    \label{fig:mesh0}
\end{figure}

@@

We plotted a Facet Grid of the returns and CPI to see further the first visual relation between the asset classes and CPI (figure \ref{fig:mesh6}). Especially the commodities and the energy sectors show a robust positive relationship according to these plots. The gold futures (GC=F) and foreign currencies (FXE) plots are interesting to see. The relationship is generally strongly positive up to point 0.4 in the CPI data. Some data points between 0.4 and 0.6 lie, which pulls the linear relationship downwards. Because it's a linear plot, the fewer points after 0.4 than before 0.4 skew the relation downwards, if one corrects for this, the relationship between the asset class and CPI would be strongly positive.

@@

\subsection{The Optimization Problem}

@@

Minimise or maximise \( f(x) \) with subject to:
\begin{itemize}
    \item \( c_j(x) &= 0, \quad j = 1, \ldots, m \text{ (equality constraints)} \)
    \item \( c_k(x) &\geq 0, \quad k = 1, \ldots, n \text{ (inequality constraints)} \)
\end{itemize}

@@

\begin{table}[H]
\centering
\begin{tabular}{ | m{7cm} | m{4cm} | }
 \hline
 \textbf{Asset Class} & \textbf{Weight} \\
 \hline
 GS10: 10-Year Treasury Constant Maturity Rate & 5\% \\ 
 \hline
 GS30: 30-Year Treasury Constant Maturity Rate & 5\% \\
 \hline
 DCOILWTICO: Crude Oil Prices: West Texas Intermediate (WTI) & 5\% \\ 
 \hline
 XLE: Energy Stocks (Energy Select Sector SPDR Fund) & 15\% \\ 
 \hline
 GC=F: Gold Futures & 30\% \\ 
 \hline
 FXE: Foreign Currencies (CurrencyShares Euro Trust) & 5\% \\
 \hline
 DBC: Commodities (Invesco DB Commodity Index Tracking Fund) & 5\% \\ 
 \hline
 BTC-USD: Cryptocurrencies (Bitcoin) & 30\% \\ 
 \hline
\end{tabular}
 \label{table:tab1}
 \caption{Optimised Portfolio Composition.}
\end{table}

@@

\begin{table}[H]
\centering
\begin{tabular}{ | m{7cm} | m{4cm} | }
 \hline
 \textbf{Measure} & \textbf{Result} \\
 \hline
 Expected Annual Return & 9.82\% \\ 
 \hline
 Risk (Annual standard deviation) & 19.96\% \\
 \hline
 Sharpe Ratio 0.4922 & 0.4922 \\ 
 \hline
\end{tabular}
 \label{table:tab2}
 \caption{KPIs.}
\end{table}

@@

We have seen that a portfolio can be constructed which considers the relationship between various asset classes and inflation. This portfolio, which is optimised according to Markowitz, significantly outperforms a simple, equally weighted portfolio. This analysis was done over a horizon of 9 years, including an economy with low inflation and a recent one with a high inflationary environment. \\

>>>>

\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage[left=3cm, right=3cm, bottom=3cm, top=3cm]{geometry}	
\usepackage{setspace}
\onehalfspacing
\usepackage{float}
\usepackage{amsmath}
\usepackage{titling} % Allows custom title configuration

\title{Markowitz Portfolio Optimization with Inflation}
\author{Pablo Duce Cabeza \\ Sergey Mirzoev \\ Marc Tschudi}
\date{December 2023}

% Custom title page
\makeatletter
\renewcommand{\maketitle}{
    \begin{titlepage}
        \vspace*{\stretch{1}}
        \begin{center}
            \Large\textbf{\@title}
        \end{center}
        \vspace{\stretch{2}}
        \begin{center}
            \large\@author
        \end{center}
        \vspace{\stretch{3}}
        \begin{center}
            \large\@date
        \end{center}
        \vspace{\stretch{0.5}}
        \begin{center}
            \includegraphics[width=0.4\textwidth]{figure/uzh_logo_e_pos.eps}
        \end{center}
    \end{titlepage}
}
\makeatother

@@

@@

We plotted a Facet Grid of the returns and CPI to see further the first visual relation between the asset classes and CPI (figure \ref{fig:mesh6}). Especially the commodities and the energy sectors show a robust positive relationship according to these plots. The gold futures (GC=F) and foreign currencies (FXE) plots are interesting to see. The relationship is generally strongly positive up to point 0.4 in the CPI data. Some data points between 0.4 and 0.6 lie, which pulls the linear relationship downwards. Because it's a linear plot, the fewer points after 0.4 than before 0.4 skew the relation downwards; if one corrects for this, the relationship between the asset class and CPI would be strongly positive.

@@

\subsection{The Optimisation Problem}

@@

Minimise or maximise \( f(x) \) subject to:
\begin{itemize}
    \item \( c_j(x) = 0, \quad j = 1, \ldots, m \) (equality constraints)
    \item \( c_k(x) \geq 0, \quad k = 1, \ldots, n \) (inequality constraints)
\end{itemize}

@@

\begin{table}[H]
\centering
\begin{tabular}{ | c | c | }
 \hline
 \textbf{Asset Class} & \textbf{Weight} \\
 \hline
 GS10: 10-Year Treasury Constant Maturity Rate & 5\% \\ 
 \hline
 GS30: 30-Year Treasury Constant Maturity Rate & 5\% \\
 \hline
 DCOILWTICO: Crude Oil Prices: West Texas Intermediate (WTI) & 5\% \\ 
 \hline
 XLE: Energy Stocks (Energy Select Sector SPDR Fund) & 15\% \\ 
 \hline
 GC=F: Gold Futures & 30\% \\ 
 \hline
 FXE: Foreign Currencies (CurrencyShares Euro Trust) & 5\% \\
 \hline
 DBC: Commodities (Invesco DB Commodity Index Tracking Fund) & 5\% \\ 
 \hline
 BTC-USD: Cryptocurrencies (Bitcoin) & 30\% \\ 
 \hline
\end{tabular}
 \label{table:tab1}
 \caption{Optimised Portfolio Composition.}
\end{table}

@@

\begin{table}[H]
\centering
\begin{tabular}{ | c | c | }
 \hline
 \textbf{Measure} & \textbf{Result} \\
 \hline
 Expected Annual Return & 9.82\% \\ 
 \hline
 Risk (Annual standard deviation) & 19.96\% \\
 \hline
 Sharpe Ratio 0.4922 & 0.4922 \\ 
 \hline
\end{tabular}
 \label{table:tab2}
 \caption{KPIs.}
\end{table}

@@

We have seen that a portfolio can be constructed considering the relationship between various asset classes and inflation. This portfolio, which is optimised according to Markowitz, significantly outperforms a simple, equally weighted portfolio. This analysis was done over a horizon of 9 years, including an economy with low inflation and a recent one with a high inflationary environment. \\

### 2023-12-18 paper/presentation.tex
Latest paper version
[LATEX] Fix compilation errors in paper and presentation

\documentclass{beamer}
\mode<beamer>{\usetheme{Madrid}}
\usepackage{graphicx} % Required for inserting images
\usepackage{eso-pic}
\usepackage[absolute,overlay]{textpos}
\setbeamertemplate{section in toc}[sections numbered]
\setbeamertemplate{subsection in toc}[ball unnumbered]
\setbeamertemplate{footline}
{
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
    \usebeamerfont{title in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.25ex,dp=1ex,center]{date in head/foot}%
    \usebeamerfont{date in head/foot}\insertshortdate{}\hspace*{2em}
    \insertframenumber{} / \inserttotalframenumber\hspace*{2ex} 
    \hspace*{6ex}
  \end{beamercolorbox}}%
  \vskip0pt%
}

@@

\title{Markowitz Portfolio Optimization with Inflation}
\author{Pablo Duce Cabeza, Djordje Nikolic, Sergey Mirzoev, Marc Tschudi}
\date{December 2023}

@@

\begin{frame}{Abstract}
  \begin{itemize}
    \item Investigate the relationship between various asset classes and inflation.
    \item Analyse CPI, Oil, Gold, Bonds, Real Estate, Bitcoin, and S\&P 500.
    \item Develop a portfolio optimized for inflation protection.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Introduction}
  \begin{itemize}
    \item Hypothesis: Certain assets have a higher correlation with inflation and therefore offer a natural hedge against inflation.
    \item Assets considered: TIPS, Gold, Real Estate, Oil, Stocks, Bonds, Bitcoin.
    \item Analysis involves exploring the relationship between inflation and selected assets.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Results: Efficient Frontier}
The optimized portfolio lies on the efficient frontier:
  \begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{paper/figure/Optimal_PF.pdf}
    \caption{Efficient Frontier}
  \end{figure}
\end{frame}

\begin{frame}{Results: Performance and Benchmark}
The optimized portfolio adjusted for the CPI outperforms the equally weighted CPI adjusted portfolio:
  \begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{paper/figure/PNL.pdf}
    \caption{Optimal Portfolio PnL}
  \end{figure}
\end{frame}

>>>>

\documentclass{beamer}
\mode<beamer>{\usetheme{Madrid}}
\usepackage{graphicx} % Required for inserting images
\usepackage{eso-pic}
\usepackage[absolute,overlay]{textpos}

\setbeamertemplate{section in toc}[sections numbered]
\setbeamertemplate{subsection in toc}[ball unnumbered]
\setbeamertemplate{footline}
{
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
    \usebeamerfont{title in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.25ex,dp=1ex,center]{date in head/foot}%
    \usebeamerfont{date in head/foot}\insertshortdate{}\hspace*{2em}
    \insertframenumber{} / \inserttotalframenumber\hspace*{2ex} 
    \hspace*{6ex}
  \end{beamercolorbox}}%
  \vskip0pt%
}

@@

\title{Markowitz Portfolio Optimization with Inflation}
\author{Pablo Duce Cabeza, Sergey Mirzoev, Marc Tschudi}
\date{December 2023}

@@

\begin{frame}{Abstract}
  \begin{itemize}
    \item Investigate the relationship between various asset classes and inflation.
    \item Analyse CPI, Oil, Gold, Bonds, Real Estate, Bitcoin, and S\&P 500.
    \item Develop a portfolio optimised for inflation protection.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Introduction}
  \begin{itemize}
    \item Hypothesis: Certain assets have a higher correlation with inflation and, therefore, offer a natural hedge against inflation.
    \item Assets considered: TIPS, Gold, Real Estate, Oil, Stocks, Bonds, Bitcoin.
    \item Analysis involves exploring the relationship between inflation and selected assets.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Results: KPIs}
The optimised portfolio exhibits the following KPIs:
\begin{table}
    \centering
    \begin{tabular}{|c|c|}
        \hline
        \textbf{Measure} & \textbf{Result} \\ \hline % Header row
        Expected Annual Return & 9.82\% \\ \hline % Data rows with horizontal lines
        Risk (Annual standard deviation) & 19.96\% \\ \hline
        Sharpe Ratio & 0.4922 \\ \hline
    \end{tabular}
    \caption{KPIs}
    \label{table:tab1}
\end{table}
\end{frame}

\begin{frame}{Results: Efficient Frontier}
The optimised portfolio lies on the efficient frontier:
  \begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{paper/figure/Optimal_PF.pdf}
    \caption{Efficient Frontier}
  \end{figure}
\end{frame}

\begin{frame}{Results: Performance and Benchmark}
The optimised portfolio adjusted for the CPI outperforms the equally weighted CPI-adjusted portfolio:
  \begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{paper/figure/PNL.pdf}
    \caption{Optimal Portfolio PnL}
  \end{figure}
\end{frame}
